A GPU Implementation of \gls{MVDR} for an Active Sonar System

### 2012-05-07 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:07	14:38:49	(Mon 7.May)

Modern phased array active sonar systems are often limited by their lateral resolution and \gls{SNR}, which is due to limited lateral extent of the array the spatial extent of the array sample the extent of the array is too small compared to the wavelength of the received signal, and while it is possible to trade resolution for improved \gls{SNR} by applying various weightsets to the array channels, one always end up with a compromise between the two. By introducing adaptive techniques, however, it is possible to resolve lateral details that would otherwise have been missed.

The \gls{MVDR} (or Capon) beamformer is one such technique. While ensuring unit gain in the look direction, it computes the set of weights that minimizes the power accumulated by the array from other directions. Whenever the distribution of noise and interference is primarily focused around specific angles, an array response with a high degree of suppression at these angles are chosen, resulting in an image with much improved lateral detail level.

However, despite its inherent potential, the \gls{MVDR} beamformer has yet to see widespread adoption in the active sonar community. There may be several reasons for this. First, the method is not inherently robust, and will suffer from a phenomena called signal cancellation in active systems. Second, in its original form, the computational complexity is cubic with the number of channels, O(M$^3$), while \gls{DAS} is only at O($M$).

Perhaps surprisingly, we will show that the suggested means for making the \gls{MVDR} behave well in active systems will also make it less computationally complex. Furthermore, by implementing it on a modern \gls{GPU} we have gained a speedup of 2-3 orders of magnitude compared to optimized single thread implementations in C and Matlab.

The feasibility of using the \gls{MVDR} beamformer in active sonar imaging has been demonstrated in e.g. \cite{Blomberg2012a,Blomberg2011,Dursun2009}.
E.g.

GPU ultrasound (\cite{So2011,Chen2011})

Complex data, all robustification techniques.

\begin{itemize}
\item Adaptive beamformer's potential lies in its ability to suppress interference power
\item Why adaptive beamformers struggle in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive. Constraints must be applied in one way or another - parameters must be tuned.
\item CUDA?
\end{itemize}

REFERENCES!!! (all over..)
\todo{\gls{MVDR} or MVDR?}
% - Who's done stuff on GPU beamforming before?
% 
% - Complex data
% -

@@

Beamforming is a process well suited for acceleration by parallelisation, because each pixel can be processed idependently.

>>>>

Modern phased array active sonar systems are often limited by their lateral resolution and \gls{SNR}. This is due to the limited lateral extent of the array, and while it is possible to trade resolution for improved \gls{SNR} by applying various weightsets to the array channels, one always end up with a compromise between the two. By introducing adaptive techniques, however, it is possible to resolve lateral details that would otherwise have been missed.

The \gls{MVDR} (or Capon) beamformer is one such technique. While ensuring unit gain in the look direction, it computes the set of weights that minimizes the power accumulated by the array from other directions. This ability to suppress interference power improves the image detail level and \gls{SNR}, and the effect can be dramatic if the noise and interference is primarily focused around specific angles.The feasibility of using the \gls{MVDR} beamformer in active sonar imaging has been demonstrated in e.g. \cite{Blomberg2012a,Blomberg2011,Dursun2009}.

However, despite its inherent potential, the \gls{MVDR} beamformer has yet to see widespread adoption in the active sonar community. There may be several reasons for this. First, the method is not inherently robust, and will suffer from a phenomena called signal cancellation in active systems. Second, in its original form, the computational complexity is cubic with the number of channels, O(M$^3$), while \gls{DAS} is only at O($M$). This makes it impractical in many situations, and even infeasible in high channel systems. \todo{mention disadvantage of tuning parameters?}

We will show that the suggested means for making the \gls{MVDR} behave well in active systems will also make it less computationally complex. Furthermore, by implementing it on a modern \gls{GPU} we have gained a speedup of around 2 orders of magnitude compared to an optimized single thread implementations in C.

Yay or nay? $\Rightarrow$ GPU ultrasound (\cite{So2011,Chen2011}). Complex data, all robustification techniques.

@@

Beamforming is a process well suited for acceleration by parallelisation, because each pixel can be processed idependently. \todo{blend it in}
